### 2020-11-09 docs/manuscript/PLOS-MSc-Thesis.tex
update Genetic algorithm section to describe Markov chain algorithm

Our understanding of how GRNs process information has increased thanks to the
concept and theory of network motifs, which are defined as patterns of
interconnections occurring in networks at numbers that are significantly higher
than those in randomized networks \cite{Milo2002}. This novel approach, along
with mathematical modeling and computational methods have been shown to be a
powerful tool to study GRNs and the predictions of studies using these methods
have been validated by experimental studies
\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and some of them 
have used synthetic GRNs to produce artificial cell differentiation in bacteria
cell cultures~\cite{Basu2005}.

@@

\begin{enumerate}
 \item{\bf Generate a set of 200 random GRNs.}

 \item{\bf Numerically solve the dynamical system for each GRN from t=0 until 
 t=500.} We integrated the differential equation using the function NDSolve from
 Wolfram Mathematica 11 with the option “EquationSimplification” and the 
 “Residual” simplification method.

 \item{\bf Evaluate fitness by comparing the phenotypes with the optimal 
 phenotype (Fig~\ref{fig:exp-pattern}).} In order to compute the fitness 
 function for each GRN we used three different filters. The first filter 
 assesses whether the expression profile of the output node reaches a 
 quasi-steady state:

@@

 \item{\bf Select GRNs with a probability proportional to its fitness.}

 \item{\bf Recombine and mutate GRNs.} Mutations included changes in the
 numerical values of degradation and diffusion parameters and interaction
 strengths, and addition and removal of interactions among genes and between the
 genes and the morphogen.
\end{enumerate}

Steps 2 to 5 were iterated 1000 times, resulting in a set of 2061 GRNs with a
striped pattern of gene expression in at least one gene.

>>>>

Our understanding of how GRNs process information has increased thanks to the
concept and theory of network motifs, which are defined as patterns of
interconnections occurring in networks at numbers that are significantly higher
than those in randomized networks \cite{Milo2002}. This novel approach, along
with mathematical modeling and computational methods have been shown to be a
powerful tool to study GRNs and the predictions of studies using these methods
have been validated by experimental studies
\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and some of them 
have even used synthetic GRNs to produce artificial cell differentiation~\cite{Basu2005} 
and circadian gene expression~\cite{atkinson_development_2003}.

@@

\begin{enumerate}
 \item{\bf Generate a random GRN.}

 \item{\bf{Numerically solve the dynamical system for the GRN from t=0 until 
t=500.} We integrated the differential equation using the function NDSolve from 
Wolfram Mathematica 11 with the option “EquationSimplification” and the 
“Residual” simplification method.}

 \item{\bf Evaluate fitness by comparing the GRN phenotype with the optimal 
 phenotype (Fig~\ref{fig:exp-pattern}).} In order to compute the fitness 
 function for each GRN we used three different filters. The first filter 
 assesses whether the expression profile of the output node reaches a 
 quasi-steady state:

@@

 \item{\bf Mutate GRN network parameters or gene product diffusion or 
degradation rates randomly and evaluate fitness again.} If fitness of the 
mutated GRN is greater than that of the previous GRN, save GRN and continue 
mutating.

 \item {\bf Mutate GRNs randomly and evaluate fitness again.} Mutations 
included changes in the  numerical values of degradation and diffusion 
parameters and interaction strengths, and addition and removal of interactions 
among genes and between the genes and the morphogen. If fitness of the mutated 
GRN was greater or equal than that of the previous GRN, mutated GRN was saved 
and the cycle continued from step 2. This cycle was repeated 5000000 times and 
GRNs with a fitness greater or equal than 0.95 were saved into a file.

\end{enumerate}

Steps 1 to 5 were iterated 500 times, resulting in a set of 2061 GRNs with a 
striped pattern of gene expression in at least one gene.
